---
title: "The Best and Worst Places to Remote Work"
author: "scurry"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    theme: sandstone
    number_sections: FALSE
    toc_float: TRUE
    toc_depth: 2
    code_folding: hide
bibliography: references.bib
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(tigris_use_cache = TRUE)
setwd("C:/LocalDocuments/MUSA6310c/Assignment02/Remote_Work")

library(dplyr)
library(ggplot2)
library(kableExtra)
library(knitr)
library(scales)
library(sf)
library(tidycensus)
library(tigris)

us <- states() %>% filter(as.numeric(REGION) <= 3)
water <- st_as_sfc(st_bbox(us))
pacounty <- counties(state = 42) %>% select(NAME)

acs_variable_list.2022 <- load_variables(2022, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)

#pawage <- read.csv("LivingWage_PA2024.csv") %>% select(COUNTY, X2A2W0C) %>% rename("NAME" = "COUNTY")
#pawage <- merge(pacounty, pawage, by = "NAME", all.x = TRUE)

#pals_24 <- read.csv("LivingSalary_PA2024.csv")

#pals_2a1w0c_24 <- pals_24 %>% select(COUNTY, X2A1W0C) %>% rename("NAME" = "COUNTY") #individual income
#pals_2a2w0c_24 <- pals_24 %>% select(COUNTY, X2A2W0C) %>% rename("NAME" = "COUNTY") #household income
#pals_2a2w1c_24 <- pals_24 %>% select(COUNTY, X2A2W1C) %>% rename("NAME" = "COUNTY") #household income

#pals_2a1w0c_24 <- merge(pacounty, pasal, by = "NAME", all.x = TRUE)

msa_mhi_22 <- get_acs(geography = "metropolitan statistical area/micropolitan statistical area",
                     year = 2022, 
                     variables = c(pop = "B01003_001", mhi = "B19013_001"),#median household income
                     geometry = TRUE,
                     output = "wide"
) 

msa_mhi_22_cnt <- msa_mhi_22 %>%
  arrange(desc(popE)) %>%
  select(NAME, mhiE) %>%
  rename("Metropolitan Statistical Area" = "NAME") %>%
  mutate("Median Household Income" = scales::dollar(mhiE)) %>%
  select(-mhiE) %>%
  st_drop_geometry() %>%
  head(8)

msa_mhi_22_cnt <- msa_mhi_22_cnt %>%
  mutate(ID = c(1:8), .before = "Metropolitan Statistical Area")

```
Remote workers are often looking for jobs based in places where they can make a high 
income. In order to maximize their paycheck, they're looking for places where
they can rent or own housing where there is a low cost of living - 
so where should they look for a job and where should they live? The top most 
populated Metropolitan Statistical Areas in the United States don't 
necessarily rank the same way in terms of how much they're paying out. 

```{r size_mhi_msa, echo=FALSE}

msa_mhi_22_cnt %>%
  kbl(col.names = c(" ", "Metropolitan Statistical Area", "Median Household Income")) %>%
  kable_material(c("triped", "hover"))

```

```{r ugh, eval=FALSE, include=FALSE}
us <- get_acs(geography = "state",
                     year = 2022, 
                     variables = c(pop = "B01003_001"),
                     geometry = TRUE,
                     output = "wide"
) %>%
  select(GEOID) %>%
#  filter(as.numeric(GEOID) <= 50) %>%
  mutate(region = case_when((GEOID = "09" | GEOID = "23" | GEOID = "25" | GEOID = "33" | GEOID = "44" | GEOID = "50" | GEOID = "34" | GEOID = "36" | GEOID = "42") ~ "Northeast",
                            (GEOID = "18" | GEOID = "17" | GEOID = "26" | GEOID = "39" | 
                              GEOID = "55" | GEOID = "19" | GEOID = "20" | GEOID = "27" |
                              GEOID = "29" | GEOID = "31" | GEOID = "38" | GEOID = "46") ~ 
                              Midwest,
                            (GEOID = "10" | GEOID = "11" | GEOID = "12" | GEOID = "13" |
                              GEOID = "24" | GEOID = "37" | GEOID = "45" | GEOID = "51" |
                              GEOID = "54" | GEOID = "01" | GEOID = "21" | GEOID = "28" |
                              GEOID = "47" | GEOID = "05" | GEOID = "22" | GEOID = "40" |
                              GEOID = "48") ~ South,
                            (GEOID = "04" | GEOID = "08" | GEOID = "16" | GEOID = "35" |
                              GEOID = "30" | GEOID = "49" | GEOID = "32" | GEOID = "56" |
                              GEOID = "02" | GEOID = "06" | GEOID = "15" | GEOID = "41" |
                              GEOID = "53") ~ West))
  
  plot(msa_mhi_22)
```

If you work remotely out of Philadelphia, where in the Northeastern region can you
live comfortably? 

In 2022, the Median Household Income in the Metropolitan Statistical Area of 
Philadelphia, PA was $85,555. Compared to the MHI of other MSAs in the United 
States,

```{r msa_mhi_data, message=FALSE, warning=FALSE, include=FALSE}

phl_mhi_22 <- msa_mhi_22 %>%
  filter(GEOID == 37980)

```


```{r msa_mhi_chart, echo=FALSE, message=FALSE, warning=FALSE}

ggplot() +
  geom_histogram(data = msa_mhi_22, aes(x = mhiE), fill = "darkolivegreen4", color = "darkolivegreen") +
  scale_x_continuous(labels = scales::label_dollar()) +
  geom_vline(data = phl_mhi_22, aes(xintercept = mhiE), color = "goldenrod", linetype = "dashed",
             size = 1) +
  geom_text(data = phl_mhi_22, aes(x = mhiE + 3000, y = 100, label = "Philadlephia MSA"),
            color = "goldenrod", angle = 90) +
  labs(title = "2022 Median Household Income",
       subtitle = "US Metropolitan Statistical Areas",
       x = "Median Household Income",
       y = " ",
#       fill = "Percentage of Philadelphia MHI",
       caption = "Source: 2022 ACS, U.S. Census Bureau") +
  theme_minimal()

```

The Philadelphia MSA's MHI is in the 93rd percentile of all MSAs. THeoretically,
securing a remote work job that was based in Philadelphia allows for a large 
amount of mobility. Across the country, a household earning that annual 
income could likely rent in most Counties that orbit an MSA wiht a lower Cost of 
Living without risk of being cost-burdened. 

A cost-burdened household pays more than 30% of its income on rent and utilities. 

The Cost of Living per County was calculated as the annual salary needed to spend 
only 30% on gross rent. In the maps below, Comfortability refers to the ratio of 
the Remote Work Base Location to the Cost of Living in the County.

```{r col_data, message=FALSE, warning=FALSE, include=FALSE}

ec_mgr_msmoc_22 <- get_acs(geography = "county",
                     year = 2022, 
                     variables = c(mgr = "B25064_001", msmoc = "B25094_001"), #GROSS RENT - SMOC
                     geometry = TRUE,
                     state = c("FL", "GA", "SC", "NC", "TN",
                               "VA", "WV", "KY", "IN", "OH",
                               "MI", "MD", "DC", "PA", "NJ",
                               "NY", "CT", "RI", "MA", "NH",
                               "VT", "ME", "DE"),
                     output = "wide"
)

# ec_msmoc_22 <- get_acs(geography = "county",
#                      year = 2022, 
#                      variables = "B25094_001",#SMOC
#                      geometry = TRUE,
#                      state = c("FL", "GA", "SC", "NC", "TN",
#                                "VA", "WV", "KY", "IN", "OH",
#                                "MI", "MD", "DC", "PA", "NJ",
#                                "NY", "CT", "RI", "MA", "NH",
#                                "VT", "ME", "DE")
#)

#take cheaper of rent vs homeowning
ec_lrs_22 <- ec_mgr_msmoc_22 %>%
  mutate(cheap = pmin(mgrE, msmocE)) %>%
  mutate(livsal = (cheap/.30)*12,
         pc_phl_mhi = (85555/livsal)) %>%
  mutate(f_phl_mhi = case_when(pc_phl_mhi >= 2.0 ~ "Twice As Comfortable",
                               pc_phl_mhi < 2.0 & pc_phl_mhi >= 1.5 ~ "Nearly Twice as Comfortable",
                               pc_phl_mhi < 1.5 & pc_phl_mhi > 1.0 ~ "More than Comfortable",
                               pc_phl_mhi == 1.0 ~ "Equally As Comfortable",
                               pc_phl_mhi < 1.0 & pc_phl_mhi >= 0.8 ~ "Uncomfortable",
                               pc_phl_mhi < 0.8 & pc_phl_mhi >= 0.5 ~ "Moderately Uncomfortable",
                               pc_phl_mhi < 0.5 & pc_phl_mhi >= 0.3 ~ "Very Uncomfortable",
                               pc_phl_mhi < 0.3 ~ "Extremely Uncomfortable"
  ))

# ec_los_22 <- ec_msmoc_22 %>%
#   select(-moe) %>%
#   mutate(livosal = (msmocE/.30)*12,
#          pc_phl_mhi = (85555/livosal)) %>%
#   mutate(f_phl_mhi = case_when(pc_phl_mhi >= 2.0 ~ "Twice As Comfortable",
#                                pc_phl_mhi < 2.0 & pc_phl_mhi >= 1.5 ~ "Nearly Twice as Comfortable",
#                                pc_phl_mhi < 1.5 & pc_phl_mhi > 1.0 ~ "More than Comfortable",
#                                pc_phl_mhi == 1.0 ~ "Equally As Comfortable",
#                                pc_phl_mhi < 1.0 & pc_phl_mhi >= 0.8 ~ "Uncomfortable",
#                                pc_phl_mhi < 0.8 & pc_phl_mhi >= 0.5 ~ "Moderately Uncomfortable",
#                                pc_phl_mhi < 0.5 & pc_phl_mhi >= 0.3 ~ "Very Uncomfortable",
#                                pc_phl_mhi < 0.3 ~ "Extremely Uncomfortable"
#  ))

ec_lrs_22$f_phl_mhi <- factor(ec_lrs_22$f_phl_mhi, levels = c("Twice As Comfortable",
                                                              "Nearly Twice as Comfortable",
                                                              "More than Comfortable",
                                                              "Equally As Comfortable",
                                                              "Uncomfortable",
                                                              "Moderately Uncomfortable",
                                                              "Very Uncomfortable",
                                                              "Extremely Uncomfortable"))

# ec_los_22$f_phl_mhi <- factor(ec_lrs_22$f_phl_mhi, levels = c("Twice As Comfortable",
#                                                               "Nearly Twice as Comfortable",
#                                                               "More than Comfortable",
#                                                               "Equally As Comfortable",
#                                                               "Uncomfortable",
#                                                               "Moderately Uncomfortable",
#                                                               "Very Uncomfortable",
#                                                               "Extremely Uncomfortable"))

```

Twice as Comfortable refers to a County where the Cost of Living is less than 
half of the Median Household Income in Philadelphia. Extremely Uncomfortable means
that the household living that County, with the salary of the Philadelphia Median 
Household Income, would be in poverty.

```{r col_map_region, echo=FALSE, message=FALSE, warning=FALSE}

ggplot() +
  geom_sf(data = water, fill = "steelblue") +
  geom_sf(data = ec_lrs_22, aes(fill = f_phl_mhi), color = "transparent") +
  scale_fill_brewer(palette = "RdYlGn", na.value = "gray", direction = -1) +
  xlim(-80, -65) +
  ylim(38, 45) +
  geom_sf(data = us, fill = "transparent", color = "steelblue", size = 8) +
  labs(title = "2022 Remote Work Bargains for Renters",
       subtitle = "Northeastern Region",
       fill = "Percentage of Philadelphia MHI",
       caption = "Source: 2022 ACS, U.S. Census Bureau") +
  theme_void() +
  theme(legend.position = "right")
```

In the Northeastern region, the only area where a household would not be comfortable 
is in the DC/Northern Virginia area. Throughout most of Pennsylvania, a househould
would be more than comfortable given the cost of living and price of rents per County.

```{r col_map_pa, echo=FALSE, message= FALSE, warning=FALSE}
ggplot() +
  geom_sf(data = water, fill = "steelblue") +
  geom_sf(data = ec_lrs_22, aes(fill = f_phl_mhi), color = "transparent") +
  scale_fill_brewer(palette = "RdYlGn", direction = -1) +
  xlim(-81, -74.5) +
  ylim(39.5, 42.5) +
  geom_sf(data = us, fill = "transparent", color = "steelblue", size = 8) +
  labs(title = "2022 Remote Work Bargains for Renters",
       subtitle = "Pennsylvania",
       fill = "Percentage of Philadelphia MHI",
       caption = "Source: 2022 ACS, U.S. Census Bureau") +
  theme_void() +
  theme(legend.position = "right")

```

